\section{Translation}

To make your django application available for a broad audience, internationalization is needed. With the translation framework of django this is possible.

\subsection{Translation definitions}
To create an internationalized version of the scorecard application, all strings need to put into translation functions. Based on these translation strings, the \emph{django-admin.py} file creates a \emph{.po} file, in which the translation takes place. First in your views (or wherever strings are created), import the translation method and enclose the strings accordingly. Look at the string in line 6.

\begin{lstlisting}[style=Python, caption=exceprt of views.py with translation, label=lst:views.py_translation]
from django.utils.translation import ugettext as _

def vote(request, pk, vote):
    course = get_object_or_404(Course, pk=pk)
    if has_already_voted(request):
        messages.add_message(request, messages.ERROR, _("You have already voted!"))
\end{lstlisting}

For translating templates, the \lstinline|{% trans "string" %}| tag is available. See how translation for the menu links is initialized in lines 5 and 6 of the following snippet.

\begin{lstlisting}[style=HTML, caption=exceprt of base.html with translation, label=lst:base.html_translation]
{% load i18n %}

<div class="navbar-collapse" id="bs-example-navbar-collapse-1">
    <ul class="nav navbar-nav">
        <li><a href="{% url 'scorecard:index' %}">{% trans "Home" %}</a></li>
        <li><a href="{% url 'scorecard:statistics' %}">{% trans "Statistics" %}</a></li>
    </ul>
</div><!-- /.navbar-collapse -->
\end{lstlisting}

\subsection{Pluralization}
Pluralization is a bit more difficult. If English is set as the first language, the plural ``s'' can be used easily with the \lstinline|pluralize| command. This is not possible for more complicated languages such as German or any than the first language used in the project. To create a pluralized version of a string, you need a counter that defines whether the singular or plural version of the string is used. In the following code, \lstinline|course.votes| is used as counter, to pluralize the votes.

\begin{lstlisting}[style=HTML, caption=exceprt of index.html with translation, label=lst:index.html_translation]
{% blocktrans count counter=course.votes %}{{ counter }} vote.{% plural %}{{ counter }} votes.{% endblocktrans %}
\end{lstlisting}

\subsection{Translation Files}
Now we have defined the strings that shall be translated. To do the actual translation, let django create a translation file. Create the folder \emph{scorecard/locale}.

From within the \emph{scorecard} directory run the following command to create a translation file for German:

\begin{lstlisting}[style=Bash, caption=Create a German translation file, label=lst:makemessages]
django-admin.py makemessages -l de
\end{lstlisting}

Now in the directory \emph{scorecard/locale}, a directory for the German language appears. Deep down in this directory there is a \emph{django.po} file. Create the translation strings within this file. You can see the plural version of the string for the votes counter in the \emph{index.html} file in line 9-14.

\begin{lstlisting}[style=HTML, caption=django.po file for German, label=lst:django.po]
#: templates/scorecard/base.html:20
msgid "Home"
msgstr "Start"

#: templates/scorecard/base.html:21
msgid "Statistics"
msgstr "Statistik"

#: templates/scorecard/index.html:10
#, python-format
msgid "%(counter)s vote."
msgid_plural "%(counter)s votes."
msgstr[0] "%(counter)s Stimme."
msgstr[1] "%(counter)s Stimmen."
\end{lstlisting}

To apply the translated messages, run:
\begin{lstlisting}[style=Bash, caption=Compile messages, label=lst:compilemessages]
django-admin.py compilemessages
\end{lstlisting}